---
title: "Illinois Income Distribution Mapping"
author: "Yoon Hong Park"
date: "July 2, 2018"
runtime: shiny
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<br><br>

This project was one of three assignments from the "Introduction to GIS and Spatial Analysis for Social Scientists" course taught by Dr. Marynia Kolak. This mapping project was an effort to present the distribution of income level accross the State of Illinois. Income data, a subset of American Community Survey, were collected from the American FactFinder, which is the data portal of the United States Census Bureau. The purpose of this project was to create multifaceted maps to compare maps from different levels of analysis, so I used two different folders for two levels of analysis, county-level and tract-level.

<br><br>

## **Basic Setup**

<br>

### **Set the current working directory**
```{r, results=TRUE, warning=FALSE}
setwd("../../2018.3 Summer/Intro to GIS/Assignment 2")
```


<br>

### **Load Libraries**
```{r, results=FALSE, warning=FALSE}
library(rgdal)
library(rgeos)
library(tmap)
library(leaflet)
library(RColorBrewer)
```



<br>

### **Load the Income data by different levels of analysis**
```{r, results=FALSE, warning=FALSE}
Census.DataCount <- read.csv("../../2018.3 Summer/Intro to GIS/Assignment 2/county-level/ACS_15_5YR_DP03_with_ann.csv")
Census.DataTract <- read.csv("../../2018.3 Summer/Intro to GIS/Assignment 2/tract-level/ACS_15_5YR_DP03_with_ann.csv")

head(Census.DataCount)
head(Census.DataTract)
```



<br>

### **Subset the Income data sets** ###
Looking up the head part of each data set, I found out that there were too many variable fields. Looking carefully at the description of each variable field, I noticed that the "HC01_VC85" field represents the median income, and the "HC01_VC86" field represents the average income and made the subset by selecting these two columns as well as the unique ID numbers. After making subsets of those two variables, I renamed those two variables by their actual contents.
```{r, results=TRUE, warning=FALSE}
var <- c("GEO.id2","HC01_VC85", "HC01_VC86")

Census.SubsetCount <- (Census.DataCount[var])
Census.SubsetTract <- (Census.DataTract[var])

names(Census.SubsetCount)[2:3]<- c("MedInc15", "AgInc15")
names(Census.SubsetTract)[2:3]<- c("MedInc15", "AgInc15")
```



<br>

### **Load the shapefile**
Before I loaded the shapefiles, I found out that there were two tracts in the tract-level shapefile which actually represented part of Michigan Lake. Therefore, I used GeoDa to tear this part off from tract-level map. There was 3123 elements in the original shapefile, but after tearing off those two tracts, it became 3123 elements.
```{r, results=TRUE, warning=FALSE}
Output.AreasCount <- readOGR("../../2018.3 Summer/Intro to GIS/Assignment 2/county-level", "IL_Counties")
Output.AreasTract <- readOGR("../../2018.3 Summer/Intro to GIS/Assignment 2/tract-level", "IL_tracts")
Output.AreasTractOrg <- readOGR("../../2018.3 Summer/Intro to GIS/Assignment 2/tract-level", "tl_2015_17_tract")

head(Output.AreasCount)
head(Output.AreasTract)
head(Output.AreasTractOrg)
```



<br><br><br>

## **Data Merging/Cleaning**

<br>

### **Merge Data with Shapefile**

By looking up the head part of both files, I noticed that the area ID numbers were written in the "GEO.id2" column in the shapefiles and in the "GEOID" column in the income data. I merged these two data sets by using these column names.
```{r, results=TRUE, warning-FALSE}
ILCounty <- merge(Output.AreasCount, Census.SubsetCount, by.x="GEOID", by.y="GEO.id2")
ILTract <- merge(Output.AreasTract, Census.SubsetTract, by.x="GEOID", by.y="GEO.id2")

head(ILCounty)
head(ILTract)
```


<br>

Before mapping, I checked if the median-income and average-income data were numeric values or not. 

```{r, results=TRUE, warning=FALSE}
str(ILCounty$MedInc15)
str(ILCounty$AgInc15)
str(ILTract$MedInc15)
str(ILTract$AgInc15)
```


<br>

I found out that they are factor-type values, so I needed to convert them into numeric values.

```{r, results=TRUE, warning=FALSE}
ILCounty$MedInc15 <- as.numeric(as.character(ILCounty$MedInc15))
ILTract$MedInc15 <- as.numeric(as.character(ILTract$MedInc15))

ILCounty$AgInc15 <- as.numeric(as.character(ILCounty$AgInc15))
ILTract$AgInc15 <- as.numeric(as.character(ILTract$AgInc15))
```



<br><br><br>

## **Mapping**

First, I Created a map for quick overview.

```{r, results=TRUE, warning=FALSE}
qtm(ILCounty, fill="MedInc15")
qtm(ILTract, fill="MedInc15")

qtm(ILCounty, fill="AgInc15")
qtm(ILTract, fill="AgInc15")
```


<br>

From this quick overview, I realized that while the county-level map seemed to be okay, the tract-level maps showed the variables were concentrated in the intervals between 0 and 100,000. With curiosity about how the variables were naturally clustered, I created another map using natural breaks. I also wanted to change the palette to shape this map aesthetically better. I added border settings since I felt the thick borders disturbed the representation of the northeastern part of Illinois, which was top-right part of the maps.

```{r, results=TRUE, warning=FALSE}
tm_shape(ILCounty) + 
  tm_fill("MedInc15", palette = "BuPu", style = "jenks") + tm_borders(lwd = 0.01, alpha = .4)
tm_shape(ILTract) + 
  tm_fill("MedInc15", palette = "BuPu", style = "jenks") + tm_borders(lwd = 0.01, alpha = .4)

tm_shape(ILCounty) + 
  tm_fill("AgInc15", palette = "BuPu", style = "jenks") + tm_borders(lwd = 0.01, alpha = .4)
tm_shape(ILTract) + 
  tm_fill("AgInc15", palette = "BuPu", style = "jenks") + tm_borders(lwd = 0.01, alpha = .4)
```


<br>

As we see the geographical "distribution" of income-level, I chose the natural break map to see how those variables were geographically clustered. I could see three clusters at top-right, center, and bottom-left part of the map. (I did not specify the number of bins for the assignment.) Then, I changed the title of legend and added a title, a legend, a compass, a scale-bar, and a source credit to each map. 

```{r, results=TRUE, warning=FALSE}
tm_shape(ILCounty) + 
  tm_fill("MedInc15", palette = "BuPu", style = "jenks", title = "Median Income ($)") + 
  tm_borders(alpha = .4) +
  tm_compass(position = c("left", "top")) + 
  tm_scale_bar(width = 0.15, position = c("left", "top")) + 
  tm_credits("U.S. Census Bureau, \n2011-2015 American Community Survey \n5-Year Estimates", align = "right", size = 0.7) +
  tm_layout(
    outer.margins = 0,
    panel.show = TRUE,
    panel.labels = "County-level Income Distribution (2015)", 
    panel.label.size = 1.15,
    panel.label.bg.color = "white",
    legend.position = c("left", "bottom"), 
    frame = FALSE)
tm_shape(ILTract) + 
  tm_fill("MedInc15", palette = "BuPu", style = "jenks", title = "Median Income ($)") + 
  tm_borders(lwd = .01, alpha = .4) + 
  tm_compass(position = c("left", "top")) + 
  tm_scale_bar(width = 0.15, position = c("left", "top")) + 
  tm_credits("U.S. Census Bureau, \n2011-2015 American Community Survey \n5-Year Estimates", align = "right", size = 0.7) +
  tm_layout(
    panel.show = TRUE,
    panel.labels = "Tract-level Income Distribution (2015)", 
    panel.label.size = 1.15,
    panel.label.bg.color = "white",
    legend.position = c("left", "bottom"), 
    frame = FALSE)

tm_shape(ILCounty) + 
  tm_fill("AgInc15", palette = "BuPu", style = "jenks", title = "Average Income ($)") + 
  tm_borders(alpha = .4) + 
  tm_compass(position = c("left", "top")) + 
  tm_scale_bar(width = 0.15, position = c("left", "top")) + 
  tm_credits("U.S. Census Bureau, \n2011-2015 American Community Survey \n5-Year Estimates", align = "right", size = 0.7) + 
  tm_layout(
    panel.show = TRUE,
    panel.labels = "County-level Income Distribution (2015)", 
    panel.label.size = 1.15,
    panel.label.bg.color = "white",
    legend.position = c("left", "bottom"), 
    frame = FALSE)
tm_shape(ILTract) + 
  tm_fill("AgInc15", palette = "BuPu", style = "jenks", title = "Average Income ($)") + 
  tm_borders(lwd = .01, alpha = .4) + 
  tm_compass(position = c("left", "top")) +
  tm_scale_bar(width = 0.15, position = c("left", "top")) + 
  tm_credits("U.S. Census Bureau, \n2011-2015 American Community Survey \n5-Year Estimates", align = "right", size = 0.7) + 
  tm_layout(
    panel.show = TRUE,
    panel.labels = "Tract-level Income Distribution (2015)", 
    panel.label.bg.color = "white",
    panel.label.size = 1.15,
    legend.position = c("left", "bottom"), 
    frame = FALSE)
```



<br>

To make multifacteted maps, I assigned a name to each mapping code and then used those names to arrange the maps. The outcomes of these codes were the final result of this assignment. 

```{r}
MedIncCount <- tm_shape(ILCounty) + 
  tm_fill("MedInc15", palette = "BuPu", style = "jenks", title = "Median Income ($)") + 
  tm_borders(alpha = .4) +
  tm_compass(position = c("left", "top")) + 
  tm_scale_bar(width = 0.15, position = c("left", "top")) + 
  tm_credits("U.S. Census Bureau, \n2011-2015 American Community Survey \n5-Year Estimates", align = "right", size = 0.7) +
  tm_layout(
    outer.margins = 0,
    panel.show = TRUE,
    panel.labels = "County-level Income Distribution (2015)", 
    panel.label.size = 1.15,
    panel.label.bg.color = "white",
    legend.position = c("left", "bottom"), 
    frame = FALSE)

MedIncTract <- tm_shape(ILTract) + 
  tm_fill("MedInc15", palette = "BuPu", style = "jenks", title = "Median Income ($)") + 
  tm_borders(lwd = 0.01, alpha = .4) + 
  tm_compass(position = c("left", "top")) + 
  tm_scale_bar(width = 0.15, position = c("left", "top")) + 
  tm_credits("U.S. Census Bureau, \n2011-2015 American Community Survey \n5-Year Estimates", align = "right", size = 0.7) +
  tm_layout(
    panel.show = TRUE,
    panel.labels = "Tract-level Income Distribution (2015)", 
    panel.label.size = 1.15,
    panel.label.bg.color = "white",
    legend.position = c("left", "bottom"), 
    frame = FALSE)

current.mode <- tmap_mode("plot") 
tmap_arrange(MedIncCount, MedIncTract)


AgIncCoun <- tm_shape(ILCounty) + 
  tm_fill("AgInc15", palette = "BuPu", style = "jenks", title = "Average Income ($)") + 
  tm_borders(alpha = .4) + 
  tm_compass(position = c("left", "top")) + 
  tm_scale_bar(width = 0.15, position = c("left", "top")) + 
  tm_credits("U.S. Census Bureau, \n2011-2015 American Community Survey \n5-Year Estimates", align = "right", size = 0.7) + 
  tm_layout(
    panel.show = TRUE,
    panel.labels = "County-level Income Distribution (2015)", 
    panel.label.size = 1.15,
    panel.label.bg.color = "white",
    legend.position = c("left", "bottom"), 
    frame = FALSE)

AgIncTrac <- tm_shape(ILTract) + 
  tm_fill("AgInc15", palette = "BuPu", style = "jenks", title = "Average Income ($)") + 
  tm_borders(lwd = .01, alpha = .4) + 
  tm_compass(position = c("left", "top")) +
  tm_scale_bar(width = 0.15, position = c("left", "top")) + 
  tm_credits("U.S. Census Bureau, \n2011-2015 American Community Survey \n5-Year Estimates", align = "right", size = 0.7) + 
  tm_layout(
    panel.show = TRUE,
    panel.labels = "Tract-level Income Distribution (2015)", 
    panel.label.bg.color = "white",
    panel.label.size = 1.15,
    legend.position = c("left", "bottom"), 
    frame = FALSE)

tmap_arrange(AgIncCoun, AgIncTrac) 
```